Structured Versus Decorated Cospans

### 2020-11-20 cospans.tex
revived old edits in section 3 and subsection A.1

$$x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2)$$}

@@

In \cite{BC2}, the first two authors introduced the symmetric monoidal double category of \emph{structured cospans} as a formalism to describe open networks. One of the main goals of this paper is to provide a monoidal double isomorphism between this double category and that of \emph{decorated cospans}, described in \cref{DecCospansDoubleCat}.  We first recall the double category of structured cospans.

\begin{thm}\label{SC}
Given categories $\X$ and $\A$ with finite colimits and $L \maps \A \to \X$ a functor preserving
finite colimits, there is a symmetric monoidal double category $_L\lCsp(\X)$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a cospan in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{displaymath}
\item a 2-morphism is a map of cospans in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{displaymath}
\end{itemize}
Composition of horizontal 1-cells and 2-cells is done using pushouts in $\X$, and the
symmetric monoidal structure is defined using binary coproducts and the initial object in $\A$ and $\X$.
\end{thm}

@@

The following theorem establishes the isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.

\begin{thm} \label{thm:equiv}
Suppose $\A$ has finite colimits and $F\maps(\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor that factors through $\Rex$ as an ordinary pseudofunctor. Then the symmetric monoidal double categories $F\lCsp$ of decorated cospans and $_L\lCsp(\inta F)$ of structured cospans are equivalent, where $L \maps \A \to \inta F$ is a left adjoint of the induced Grothendieck opfibration $U\maps \inta F \to \A$.
\end{thm}

@@

In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting.  Suppose $\A$ is a monoidal category.     Then there is a bijection between the following two things:
\begin{itemize}
\item lax monoidal structures on a pseudofunctor $F\maps\A\to\Cat$
\item monoidal structures on $\inta F$ such that $U \maps \inta F \to \A$ is strict monoidal and the tensor product in $\inta F$ preserves cocartesian liftings.  
\end{itemize}
We call the former \define{monoidal opindexed categories}, and the latter \define{monoidal opfibrations}.     If the monoidal structure of $\A$ is cocartesian, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps\A\to\bicat{MonCat}$ into the 2-category of monoidal categories, strong monoidal functors and monoidal natural transformations.  These correspondences can be enhanced to 2-equivalences:

@@

\begin{proof}
This was shown by Moeller and the third author \cite[Theorems~3.13\&4.1]{MV}. In summary, for a cocartesian base we have correspondences
\begin{gather}
\textrm{lax monoidal pseudofunctors }F\maps(\A,+)\to(\Cat,\times) \notag\\
\Updownarrow \notag\\
\textrm{monoidal opfibrations }U\maps(\X,\otimes)\to(\A,+) \notag\\
\Updownarrow \notag\\
\textrm{pseudofunctors } \tilde{F} \maps \A\to\bicat{MonCat} \notag
\end{gather}
The second equivalence was also earlier observed by Shulman \cite{Shulman2008}. In more detail, if $(\phi,\phi_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{sec:2cats}, each fibre category $\X_a=F(a)$ obtains a monoidal structure via
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\phi_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\mathbf{1}\xrightarrow{\phi_0}F(0)\xrightarrow{F(!)}F(a)
\end{equation}
where $\nabla$ is the fold map. 
\end{proof}

These correspondences further restrict to the case when the Grothendieck category is specifically cocartesian monoidal itself. In that case, opfibrations
$U \maps (\X,+)\to(\A,+)$ that strictly preserve chosen coproducts and initial object
%, and such that $+$ preserves cocartesian liftings, Christina: this should ALWAYS be the case, by universal properties!
bijectively correspond to pseudofunctors into the 2-category of cocartesian categories.  For more details, see \cite[Corollary 4.7]{MV} and the related discussion.

Finally, we can further restrict to the setting of opfibrations that also preserve pushouts, and thus all finite colimits, thanks to the following result.
%as well as Remark~4.11 therein regarding the non-necessary strictness condition. Is this obscuring or clarifying?

\begin{lem}[Hermida] \label{lem:fibrewiselimits}
Suppose $\J$ is a small category and $U\maps\X\to\A$ is an opfibration. If the base $\A$ has $\J$-colimits,
the following are equivalent:
\begin{enumerate}
 \item all fibres have $\J$-colimits, and the reindexing functors preserve them;
 \item the total category $\X$ has $\J$-colimits, and $U$ preserves them.
\end{enumerate}
\end{lem}

\begin{proof}  See \cite[Corollary~4.9]{Hermida1999}.   \end{proof}

The first part formulates the existence of colimits \emph{locally} in each fibre, and if we let $\J$ range over all finite categories it says that the corresponding pseudofunctor $F\maps\A\to\Cat$ lands in the sub-2-category $\Rex$. The second part formulates the existence of colimits \emph{globally} in the total category $\inta F$, and if we let $\J$ range over all finite
categories it says that $\X$ is finitely cocomplete and $U$ preserves all finite colimits.  Combining this with \cref{lem:MonGroth}, we deduce the following.

@@

Notice a subtlety in the above formulation. \cref{lem:MonGroth} provides a specific monoidal structure on the fibres which becomes \cref{eq:explicitstructure1} when the base $\A$ is cocartesian monoidal.   If this fibrewise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to any other cocartesian monoidal structure on the fibres, by the universal property of colimits. Therefore, we can simplify the above assumption to that in \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor. However, we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$, and that is why the correspondence in \cref{lem:MonGroth} is necessary.
DOES THIS MAKE SENSE TO ALL?? %still, we ARE going to use the formation of coproducts as above. Otherwise someone might think that the monoidal Groth story was redundant here!

On a highly related note, in what follows we are also interested in the existence of a left adjoint $L_F$ to an induced monoidal opfibration; the following result provides sufficient conditions for that.

@@

@@

%For a discussion on the unusual strict cocontinuity condition...
%\begin{lem}\label{prop:CV}\cite[Proposition~3.3]{CV}
%Suppose $\U\maps\X\to\A$ is a right adjoint `left inverse'. If $\X$ and $\A$ have chosen pushouts and initial objects and $\U$ strictly preserves them, %then $\U$ is an opfibration.
%\end{lem}

Notice that under \cref{lem:fibrewiselimits}, if the base category $\A$ has an initial object denoted $0_\A$, the above lemma is equivalent to $\X$ having an initial object $0_\X$ above $0_\A$. The way in which these two conditions give one another in this case is as follows. 
The fibrewise initial object $\bot_a$ is precisely the cocartesian lifting of $0_\X$ along the unique map $u_a\maps0_\A\to a$ in the base category
\begin{displaymath}
\xymatrix @C=.4in @R=.2in
{0_\X \ar @{.>}[d]\ar[rr]^-{\mathrm{Cocart}(0_\X,u_a)} && (u_a)_!(0_X)=:\bot_a\ar @{.>}[d] & \textrm{in }\X  \\
0_\A\ar[rr]^-{\exists!u_a} && a & \textrm{in }\A}
\end{displaymath}
Moreover, if the opfibration $U$ comes from the Grothendieck construction on a pseudofunctor $F\maps\A\to\Cat$, the reindexing functors $(u_a)_!$ of the opfibration are precisely $F(u_a)$ therefore $\bot_a=(a,F(u_a)(0_\X))$ in the Grothendieck category notation (\cref{def:GrothCat}).

Finally, if we start with a lax monoidal pseudofunctor $(F,\phi,\phi_0)$, the monoidal Grothendieck construction of \cref{lem:MonGroth} in the cocartesian case expresses $\bot_a$ as follows by \cref{eq:explicitstructure1}:
\begin{equation}\label{eq:initialinfibre}
\mathbf{1}\xrightarrow{\phi_0}F(0_\A)\xrightarrow{F(u_a)}F(a) 
\end{equation}

@@

By \cref{cor:fcocMonGroth}, when $F$ as an ordinary pseudofunctor factors through $\Rex$, the Gro\-the\-ndieck construction gives rise to a finitely cocomplete $\inta F$ such that the corresponding opfibration $U\maps\inta{F}\to\A$ preserves all finite colimits. Moreover, by \cref{prop:opfibtolari}, $U$ has a left adjoint $L\maps\A\to\inta F$ with $UL=1_\A$. Diagrammatically,
\begin{displaymath}
 F\maps\A\to\Cat\quad\mapsto\quad\begin{tikzcd}[baseline=.3]\inta F\ar[d,"U"'] \\ \A \end{tikzcd}\quad\mapsto\quad\begin{tikzcd}\A\ar[r,bend left,pos=.55,"L"]\ar[r,phantom,"\bot"description] & \inta F\ar[l,bend left,pos=.45,"U"]\end{tikzcd}
\end{displaymath}
describes the construction of the adjunction from the original $F$. Explicitly, the left adjoint picks the initial object in the finitely cocomplete fibre $L(a)=(a,\bot_a)$, also expressed as $F(u_a)\circ\\phi_0(*)$ according to \cref{eq:initialinfibre} for $(\phi,\phi_0)$ the monoidal structure of $F$.

As a left adjoint, this induced $L$ preserves all colimits that exist between the finitely cocomplete categories $\A$ and $\inta F$, so we can construct the double category of structured cospans $_L\lCsp(\inta F)$ of \cref{SC}. Its objects and vertical morphisms are those of the category $\A$ (as for $F\lCsp$ earlier), whereas horizontal 1-cells are now cospans of the form $\cspn{L(a)}{v}{L(b)}{}{}$ in the Grothendieck category $\inta F$ (\cref{def:GrothCat}). Explicitly, they consist of two pairs of morphisms 
\begin{equation}\label{eq:scsphor1cell}
 (a,\bot_a)\xrightarrow{\scalebox{0.7}{\(\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}\)}}(m,x)\xleftarrow{\scalebox{0.7}{\(\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}\)}}(b,\bot_b)
\end{equation}
with $x\in F(m)$; recall that the reindexing functors $F(i),F(o)$ preserve all finite colimits across the fibre categories.
Finally, the 2-morphisms of this double category in this context unravel as follows
\begin{equation}\label{eq:2morphismLCspintaF}
 \begin{tikzcd}[row sep=1in,column sep=1.7in,ampersand replacement=\&]
 (a,\bot_a)\ar[r,"{\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}}"]\ar[d,"{\begin{cases}f\maps a\to a' &\textrm{in }\A \\\chi_a\maps F(f)(\bot_a)\cong \bot_{a'} &\textrm{in }F(a')\end{cases}}"description] \& (m,x) \ar[d,"{\begin{cases}k\maps m\to m' &\textrm{in }\A \\ \tau\maps F(k)(x)\to x' &\textrm{in }F(m')\end{cases}}"description] \& (b,\bot_b)\ar[l,"{\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}}"']\ar[d,"{\begin{cases}g\maps b\to b' &\textrm{in }\A \\\chi_b\maps F(g)(\bot_b)\cong \bot_{b'} &\textrm{in }F(b')\end{cases}}"description] \\
 (a',\bot_{a'})\ar[r,"{\begin{cases}i'\maps a'\to m' &\textrm{in }\A \\!\maps F(i')(\bot_{a'})\to x' &\textrm{in }F(m')\end{cases}}"'] \& (m',x') \& (b',\bot_{b'})\ar[l,"{\begin{cases}o'\maps b'\to m' &\textrm{in }\A \\!\maps F(o')(\bot_{b'})\to x' &\textrm{in }F(m')\end{cases}}"]
 \end{tikzcd}
\end{equation}
where the outside vertical legs come from the definition of $L$ on arrows \cref{eq:Lonarrows}. 
Explicitly, the above square commutativities translate to the conditions $k\circ i=i'\circ f$ and $k\circ o=o'\circ g$ in $\A$, and also to
\begin{gather}\label{eq:Grothcommutativity}
 F(k\circ i)(\bot_a)\cong Fk(Fi(\bot_a))\xrightarrow{Fk(!)}Fk(x)\xrightarrow{\tau}x'= \\
 F(i'\circ f)(\bot_a)\cong Fi'(Ff(\bot_a))\xrightarrow{Fi'(\chi_a)}Fi'(\bot_{a'})\xrightarrow{!}x'\nonumber
\end{gather}
in the fibre category $F(m')$, by composition in the Grothendieck category \cref{eq:compGrothcat}. Since both morphisms emanate from the mapped initial object in $F(m')$ under the finitely cocontinuous reindexing functors, they are the unique such into $x'$ therefore this gives no extra conditions; similarly for the equality including $o,o'$.

>>>>

\begin{equation}\label{eq:tensordecoration}
x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2)

@@

In \cite{BC2}, the first two authors introduce the symmetric monoidal double category of \emph{structured cospans} as a formalism to capture open networks. One of the main goals of this paper is to provide a monoidal double isomorphism between this double category and that of \emph{decorated cospans}, described in detail in \cref{DecCospansDoubleCat}. We first recall the double category of structured cospans.

\begin{thm}\label{thm:SC}
Given categories $\X$ and $\A$ with finite colimits and $L \maps \A \to \X$ a functor preserving
finite colimits, there is a symmetric monoidal double category $_L\lCsp(\X)$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a cospan in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\end{displaymath}
\item a 2-morphism is a map of cospans in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{displaymath}
\end{itemize}
Composition of horizontal 1-cells and 2-cells is done using pushouts in $\X$, and the
symmetric monoidal structure is defined using binary coproducts and the initial object in $\A$ and $\X$: 
the tensor of horizontal 1-cells is
\begin{equation}\label{eq:tensorstuctured}
\begin{tikzcd}[column sep={.4in,between origins}, row sep=.1in]
& x &&&& x' &&&& x+x' & \\
&&&\ot&&&& = &&&& \\
L(a)\ar[uur,"i"]&& L(b)\ar[uul,"o"'] && L(a')\ar[uur,"i'"] && L(b')\ar[uul,"o'"'] && L(a+a')\ar[uur,"i+i'"] && L(b+b')\ar[uul,"o+o'"'] 
\end{tikzcd}
\end{equation}
using that $L$ preserves coproducts.
\end{thm}

@@

The following theorem establishes the isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.   (In \cite{BC} we took as objects finitely cocomplete categories with chosen finite colimits, just so we could turn these into symmetric monoidal categories without any further choices!!!  Should we do that here???) {\chris Can you tell if we can get away without doing that? Perhaps after you read up to before the proof?}

\begin{thm} \label{thm:equiv}
Suppose $\A$ has finite colimits and $F\maps(\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor that factors through $\Rex$ as an ordinary pseudofunctor. Then the symmetric monoidal double categories $F\lCsp$ of decorated cospans and $_L\lCsp(\inta F)$ of structured cospans are isomorphic, where $L \maps \A \to \inta F$ is a left adjoint of the induced Grothendieck opfibration $U\maps \inta F \to \A$.
\end{thm}

@@

In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting.  Suppose $\A$ is a monoidal category. Then there is a bijection between the following two things:
\begin{itemize}
\item lax monoidal structures on a pseudofunctor $F\maps\A\to\Cat$
\item monoidal structures on $\inta F$ such that $U \maps \inta F \to \A$ is strict monoidal and the tensor product in $\inta F$ preserves cocartesian liftings.  
\end{itemize}
We call the former \define{monoidal opindexed categories}, and the latter \define{monoidal opfibrations}. If the monoidal structure of $\A$ is cocartesian, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps\A\to\bicat{MonCat}$ into the 2-category of monoidal categories, strong monoidal functors and monoidal natural transformations.  These correspondences can be enhanced to 2-equivalences:
%-In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting: given a monoidal category $\A$ there is a bijection between \define{monoidal opindexed categories}, which are lax monoidal pseudofunctors $F\maps\A\to\Cat$, and \define{monoidal opfibrations}, which are opfibrations $U \maps \X \to \A$ where $\X$ is monoidal, $U$ is strict monoidal, and $\otimes_\X$ preserves cocartesian liftings.   This bijection sends a monoidal opindexed category $F \maps \A \to \Cat$ to its Grothendieck construction $U \maps \inta F \to \A$.  If the monoidal structure on $\A$ is in fact cocartesian, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps\A\to\bicat{MonCat}$ into the 2-category of monoidal categories, strong monoidal functors and monoidal natural transformations.   Further, all these correspondences lift to 2-equivalences:
{\chris There is a commented out passage of your earlier take here. I think I preferred your earlier more discussional version. Now having the two bullets above almost totally coincides with the three lines with arrows between them in the sketch of proof below. Feels like saying the same thing twice, whereas earlier you were really just giving brief definitions before diving in the theorem.}

@@

\begin{proof}
This was shown by Moeller and the third author \cite[Theorems~3.13\&4.1]{MV}. In summary, for a cocartesian base we have correspondences
\begin{gather}
\textrm{lax monoidal pseudofunctors }F\maps(\A,+)\to(\Cat,\times) \notag\\
\Updownarrow \notag\\
\textrm{monoidal opfibrations }U\maps(\X,\otimes)\to(\A,+) \notag\\
\Updownarrow \notag\\
\textrm{pseudofunctors } \tilde{F} \maps \A\to\bicat{MonCat} \notag
\end{gather}
The second equivalence was also earlier observed by Shulman \cite{Shulman2008}.

In more detail, if $(\phi,\phi_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{subsec:bicats}, the induced monoidal structure on the total Grothendieck category $\X=\inta F$ (\cref{def:GrothCat}) is given by
\begin{equation}\label{eq:explicitstructure2}
\Big(a,x\in F(a)\Big)\ot_{\X}\Big(b,y\in F(b)\Big)=\Big(a+b,\phi_{a,b}(x,y)\in F(a+b)\Big), \qquad I_{\X}=\Big(0_A,\phi_0\Big)
\end{equation}
Moreover, each fibre category $\X_a=F(a)$ obtains a monoidal structure via 
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\phi_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\mathbf{1}\xrightarrow{\phi_0}F(0)\xrightarrow{F(!)}F(a)
\end{equation}
where $\nabla$ is the fold map. 
\end{proof}

These 2-equivalences further restrict to the case when the Grothendieck category $(X,\otimes)$ is specifically cocartesian monoidal itself, with coproducts built up from \cref{eq:explicitstructure2}. In that case, opfibrations $(\X,+)\to(\A,+)$ that (strictly) preserve coproducts and initial object bijectively correspond to pseudofunctors into the 2-category of cocartesian categories.  For more details, see \cite[Corollary 4.7]{MV} and the related discussion.

Finally, we can further restrict to the setting of opfibrations that also preserve pushouts, and thus all finite colimits, thanks to the following more general result.

\begin{lem}[Hermida] \label{lem:fibrewiselimits}
Suppose $\J$ is a small category and $U\maps\X\to\A$ is an opfibration. If the base $\A$ has $\J$-colimits,
the following are equivalent:
\begin{enumerate}
 \item all fibres have $\J$-colimits, and the reindexing functors preserve them;
 \item the total category $\X$ has $\J$-colimits, and $U$ (strictly) preserves them.
\end{enumerate}
\end{lem}

\begin{proof}
See \cite[Corollary~4.9]{Hermida1999} as well as \cite[Remark~4.11]{Hermida1999} for the non-necessary strictness condition.
\end{proof}

The first part formulates the existence of colimits \emph{locally} in each fibre, and if we let $\J$ range over all finite categories it says that the corresponding pseudofunctor $F\maps\A\to\Cat$ lands in the sub-2-category $\Rex$. The second part formulates the existence of colimits \emph{globally} in the total category $\inta F$, and if we let $\J$ range over all finite categories it says that $\X$ is finitely cocomplete and $U$ preserves all finite colimits. As an example, suppose that $\A$ has pushouts and a Grothendieck opfibration $U\colon\inta F\to\A$ has fibrewise pushouts preserved by its reindexing functors. We can construct pushouts in $\inta F$ as follows: 
\begin{equation}\label{eq:globalpushout}
 \begin{tikzcd}[ampersand replacement=\&,column sep=1in,row sep=.5in]
\& (a+_bc,w)\ar[dd,phantom, very near start,"\upback"description] \& \\
(a,x)\ar[ur] \&\& (c,z)\ar[ul] \\
\& (b,y)\ar[ul,"{\begin{cases}f\maps b\to a &\textrm{in }\A \\k\maps F(f)y\to x &\textrm{in }F(a)\end{cases}}"description]\ar[ur,"{\begin{cases}g\maps b\to c &\textrm{in }\A \\\ell\maps F(g)y\to z &\textrm{in }F(c)\end{cases}}"description] \&
 \end{tikzcd}
\end{equation}
which is the pushout $\sigma_a\colon a\to a+_bc\leftarrow c:\sigma_c$ in $\A$ on the first variable, and $w$ is the pushout in the fibre $F(a+_bc)$
\begin{displaymath}
\begin{tikzcd}
 & w \ar[dd,phantom, very near start,"\upback"description] & \\
 F(\sigma_a)x\ar[ur] & & F(\sigma_c)z\ar[ul] \\
 & F(\sigma_a)(Ff(y))\cong F(\sigma_n)(Fg(y))\ar[ul,"F(\sigma_a)k"]\ar[ur,"F(\sigma_c)\ell"'] &
 \end{tikzcd}
\end{displaymath}
In similar ways we can construct the initial object and coproducts in a Grothendieck category.

In our case of interest, the pseudofunctor $F$ will originally be lax monoidal; combining \cref{lem:MonGroth,lem:fibrewiselimits}, we deduce the following. %Christina: I suspect that's why for composition we get isomorphism below - using pushouts in which I cannot use the monoidal structure of the functor - whereas for coproducts and initial object we get equality since they were constructed in a very specific way. Related to the subtlety comment below.

@@

Notice a subtlety in the above formulation. \cref{lem:MonGroth} provides a specific monoidal structure on the fibres which becomes \cref{eq:explicitstructure1} when the base $\A$ is cocartesian monoidal.   If this fibrewise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to any other cocartesian monoidal structure on the fibres, by the universal property of colimits. Therefore, we can simplify the above assumption to that in \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor. However, we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$, and that is why the correspondence in \cref{lem:MonGroth} is necessary.
DOES THIS MAKE SENSE TO ALL??

On a highly related note, in what follows we are also interested in the existence of a left adjoint $L$ to an induced monoidal opfibration; the following result provides sufficient conditions for that.

@@

%Notice that under \cref{lem:fibrewiselimits}, if the base category $\A$ has an initial object denoted $0_\A$, the above lemma is equivalent to $\X$ having an initial object $0_\X$ above $0_\A$.

@@

%Christina: I don't think the following is needed.
%Notice that under \cref{lem:fibrewiselimits}, if the base category $\A$ has an initial object denoted $0_\A$, the above lemma is equivalent to $\X$ having an initial object $0_\X$ above $0_\A$. The way in which these two conditions give one another in this case is as follows. 
%The fibrewise initial object $\bot_a$ is precisely the cocartesian lifting of $0_\X$ along the unique map $u_a\maps0_\A\to a$ in the base category
%\begin{displaymath}
%\xymatrix @C=.4in @R=.2in
%{0_\X \ar @{.>}[d]\ar[rr]^-{\mathrm{Cocart}(0_\X,u_a)} && (u_a)_!(0_X)=:\bot_a\ar @{.>}[d] & \textrm{in }\X  \\
%0_\A\ar[rr]^-{\exists!u_a} && a & \textrm{in }\A}
%\end{displaymath}
%Moreover, if the opfibration $U$ comes from the Grothendieck construction on a pseudofunctor $F\maps\A\to\Cat$, the reindexing functors $(u_a)_!$ of the %opfibration are precisely $F(u_a)$ therefore $\bot_a=(a,F(u_a)(0_\X))$ in the Grothendieck category notation (\cref{def:GrothCat}).

%Finally, if we start with a lax monoidal pseudofunctor $(F,\phi,\phi_0)$, the monoidal Grothendieck construction of \cref{lem:MonGroth} in the cocartesian case expresses $\bot_a$ as follows by \cref{eq:explicitstructure1}:
%\begin{equation}\label{eq:initialinfibre}
%\mathbf{1}\xrightarrow{\phi_0}F(0_\A)\xrightarrow{F(u_a)}F(a) 
%\end{equation}

@@

By \cref{cor:fcocMonGroth}, when $F$ as an ordinary pseudofunctor factors through $\Rex$, the Gro\-the\-ndieck construction gives rise to a finitely cocomplete $\inta F$ such that the corresponding opfibration $U\maps\inta{F}\to\A$ preserves all finite colimits. Moreover, by \cref{prop:opfibtolari}, $U$ has a left adjoint $L\maps\A\to\inta F$ with $UL=1_\A$. Diagrammatically,
\begin{displaymath}
 F\maps\A\to\Cat\quad\mapsto\quad\begin{tikzcd}[baseline=.3]\inta F\ar[d,"U"'] \\ \A \end{tikzcd}\quad\mapsto\quad\begin{tikzcd}\A\ar[r,bend left,pos=.55,"L"]\ar[r,phantom,"\bot"description] & \inta F\ar[l,bend left,pos=.45,"U"]\end{tikzcd}
\end{displaymath}
describes the construction of the adjunction from the original $F$. Explicitly, the left adjoint picks the initial object in the finitely cocomplete fibre $L(a)=(a,\bot_a)$, also expressed as $F(!_a)\circ\phi_0(*)$ according to \cref{eq:explicitstructure1} for $(\phi,\phi_0)$ the monoidal structure of $F$. %Christina: If each adjoint picks one isomorphic initial object, can we choose the one with that choice? Well they are all isomorphic and that could help with the monoidal structure later...

As a left adjoint, this induced $L$ preserves all colimits that exist between the finitely cocomplete categories $\A$ and $\inta F$, so we can construct the double category of structured cospans $_L\lCsp(\inta F)$ of \cref{thm:SC}. Its objects and vertical morphisms are those of the category $\A$ (as for $F\lCsp$ earlier), whereas horizontal 1-cells are now cospans of the form $\cspn{L(a)}{v}{L(b)}{}{}$ in the Grothendieck category $\inta F$ (\cref{def:GrothCat}). Explicitly, they consist of two pairs of morphisms 
\begin{equation}\label{eq:scsphor1cell}
 (a,\bot_a)\xrightarrow{\scalebox{0.7}{\(\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}\)}}(m,x)\xleftarrow{\scalebox{0.7}{\(\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}\)}}(b,\bot_b)
\end{equation}
with $x\in F(m)$; recall that the reindexing functors $F(i),F(o)$ preserve all finite colimits across the fibre categories.
Finally, the 2-morphisms of this double category in this context unravel as follows
\begin{equation}\label{eq:2morphismLCspintaF}
 \begin{tikzcd}[row sep=1in,column sep=1.7in,ampersand replacement=\&]
 (a,\bot_a)\ar[r,"{\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}}"]\ar[d,"{\begin{cases}f\maps a\to a' &\textrm{in }\A \\\chi_a\maps F(f)(\bot_a)\cong \bot_{a'} &\textrm{in }F(a')\end{cases}}"description] \& (m,x) \ar[d,"{\begin{cases}k\maps m\to m' &\textrm{in }\A \\ \tau\maps F(k)(x)\to x' &\textrm{in }F(m')\end{cases}}"description] \& (b,\bot_b)\ar[l,"{\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}}"']\ar[d,"{\begin{cases}g\maps b\to b' &\textrm{in }\A \\\chi_b\maps F(g)(\bot_b)\cong \bot_{b'} &\textrm{in }F(b')\end{cases}}"description] \\
 (a',\bot_{a'})\ar[r,"{\begin{cases}i'\maps a'\to m' &\textrm{in }\A \\!\maps F(i')(\bot_{a'})\to x' &\textrm{in }F(m')\end{cases}}"'] \& (m',x') \& (b',\bot_{b'})\ar[l,"{\begin{cases}o'\maps b'\to m' &\textrm{in }\A \\!\maps F(o')(\bot_{b'})\to x' &\textrm{in }F(m')\end{cases}}"]
 \end{tikzcd}
\end{equation}
where the outside vertical legs come from the definition of $L$ on arrows \cref{eq:Lonarrows}. 
Explicitly, the above square commutativities translate to the conditions $k\circ i=i'\circ f$ and $k\circ o=o'\circ g$ in $\A$, and also to
\begin{gather}\label{eq:Grothcommutativity}
 F(k\circ i)(\bot_a)\cong Fk(Fi(\bot_a))\xrightarrow{Fk(!)}Fk(x)\xrightarrow{\tau}x'= \\
 F(i'\circ f)(\bot_a)\cong Fi'(Ff(\bot_a))\xrightarrow{Fi'(\chi_a)}Fi'(\bot_{a'})\xrightarrow{!}x'\nonumber
\end{gather}
in the fibre category $F(m')$, by composition in the Grothendieck category \cref{eq:compGrothcat}. Since both morphisms emanate from the mapped initial object in $F(m')$ under the finitely cocontinuous reindexing functors, they are the unique such into $x'$ therefore this gives no extra conditions; similarly for the equality including $o,o'$.
